Financialisation of Commodity Markets

### 2024-09-26 docs/paper.qmd

In a similar vein, the average market factor beta for the entire cross-section drops down to 0.94 (@tbl-regressions-index), below the level observed during financialization. The decline seems driven by the metals sector, particularly the GB traded segment, with the figures for the US and GB traded assets at 0.91 and 0.72, down from 1.12 and 1.18, respectively. Meanwhile, the pricing performance of the market factor also retreats to below financialisation levels with the average $R^{2}$ for the entire cross-section dropping down to 14.7% (@tbl-regressions-factors). All sectors seem affected with agriculturals, energy and metals all seeing their average $R^{2}$ decline below the levels observed during the financialisation period albeit remaining above 10% for the GB traded metals against less than 2% pre-financialisation. Similarly, the performance for both the CHP and term structure factors falls back to levels below those attained over financialisation while the gap between their long and short legs remains narrower than observed during that period. Taken together, these results, along with those from the crisis period, suggest that the exogenous shock of the financial crisis, combined with the subsequent accommodative monetary policy, may have triggered a reversion toward pre-financialization fundamentals in the commodity futures complex; an issue that, to our knowledge, has not been widely noted in the existing literature.  
Meanwhile the regime based analysis is also revealing starting with the average correlations are higher during phases of **contango** and the pattern stronger for the metals assets both US and GB traded. Similarly, for these the pricing performance of the market factor is considerably better during phase of **contango** in contrast with the previous periods that showed no substantial difference between phase of **contango** and **backwardation**. The figures for the US and GB traded segments of the sector are 23% (**contango**) vs 11.5% (**backwardation**) against 36.2% vs 43% during financialisation and 14.4% (**contango**) vs 6.6% (**backwardation**) against 33.5% vs 34.4% during financialisation respectively which indicate some recent changes in the correlation structure within the sector. Taken together with the decline in interdependence in the commodity sector at large documented above, this observation could form a valuable basis for future research.

>>>>

In a similar vein, the average market factor beta for the entire cross-section drops down to 0.94 (@tbl-regressions-index), below the level observed during financialization. The decline seems driven by the metals sector, particularly the GB traded segment, with the figures for the US and GB traded assets at 0.91 and 0.72, down from 1.12 and 1.18, respectively. Meanwhile, the pricing performance of the market factor also retreats to below financialisation levels with the average $R^{2}$ for the entire cross-section dropping down to 14.7% (@tbl-regressions-factors). All sectors seem affected with agriculturals, energy and metals all seeing their average $R^{2}$ decline below the levels observed during the financialisation period albeit remaining above 10% for the GB traded metals against less than 2% pre-financialisation. Similarly, the performance for both the CHP and term structure factors falls back to levels below those attained over financialisation while the gap between their long and short legs remains narrower than observed during that period. These results, along with those from the crisis period, suggest that the exogenous shock of the financial crisis, combined with the subsequent accommodative monetary policy, may have triggered a reversion toward pre-financialization fundamentals in the commodity futures complex; an issue that, to our knowledge, has not been widely discussed in the existing literature.  
The phases analysis further documents the emerging role/dominance of the **contango** regime post-crisis with average correlations higher during those phases and the pattern stronger for the metal assets both US and GB traded. Similarly, for the latter the pricing performance of the market factor is considerably better during phase of **contango** in contrast with the previous periods that showed no substantial difference between phase of **contango** and **backwardation**. The figures for the US and GB traded segments of the sector are 23% (**contango**) vs 11.5% (**backwardation**) against 36.2% vs 43% during financialisation and 14.4% (**contango**) vs 6.6% (**backwardation**) against 33.5% vs 34.4% during financialisation respectively. Taken together, these results indicate some recent changes in the correlation structure within the sector and along with the decline in interdependence in the commodity sector at large documented above, this observation could form a valuable basis for future research.   
These findings further providence additional evidence that the interdependence effects of financialization seems to have been strongest for the entire cross-section of the metals sector with a lesser impact for the other sectors.
